---
title: "HMW 7- Data 605"
author: "OMER OZEREN"
output:
  word_document:
    toc: yes
    toc_depth: '5'
  html_document:
    highlight: tango
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, collapse = T)
```

## Problem 1

Let $X_1, X_2,...,X_n$ be $n$ mutually independent random variables, each of
which is uniformly distributed on the integers from $1$ to $k$. Let $Y$ denote the
minimum of the $X_i$'s. Find the distribution of $Y$. 

### SOLUTION

Number of possible combinations of $X_i$'s is $k^n$ (choosing $n$ values out of $k$ options with replacement).

Consider number of combinations with at least one $1$. It is equal to all combinations ($k^n$) minus all combinations with values between $2$ and $k$ ($(k-1)^n$). So $P(Y=1) = \frac{k^n-(k-1)^n}{k^n}$.

Consider number of combinations with at least one $2$ and no $1$. It is euqal to all combinations ($k^n$) minus all combinations with at least one $1$ (see above: $k^n-(k-1)^n$) and minus all combinations with values between $3$ and $k$ ($(k-2)^n$). So $P(Y=2) = \frac{k^n-(k^n-(k-1)^n)-(k-2)^n}{k^n}= \frac{k^n-k^n+(k-1)^n-(k-2)^n}{k^n}= \frac{(k-1)^n-(k-2)^n}{k^n}$.

Similarly considering combinations without $1$ or $2$ and with at least one $3$, 

$$
\begin{split}
P(Y=3) &=\frac{k^n - (k^n-(k-1)^n)-((k-1)^n-(k-2)^n)-(k-3)^n}{k^n}\\ 
&=\frac{k^n - k^n+(k-1)^n-(k-1)^n+(k-2)^n-(k-3)^n}{k^n}\\ 
&= \frac{(k-2)^n-(k-3)^n}{k^n}
\end{split}
$$.

More generally, we can see that $P(Y=a) = \frac{(k-a+1)^n-(k-a)^n}{k^n}$.

## Problem 2.
Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.)

### a.
What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years.)

#### Solution
Let $p$ be the probability that the machine fails, and $q = 1 - p$ the probability that it doesn't.

We're looking for the first failure (success) after 8 years, so this is a geometric distribution. Since we're only expecting one failure in 10 years, $p = 0.1, \ q = 0.9$.

$P(X > 8) = 1 - P(X \leq 8)$

$1 - (1 - q^{i+1}) = q^{i+1} = 0.9^{9} =$ `r round(0.9^9, 4)`.

$E[X] = \frac{1}{p} = \frac{1}{0.1} =$ `r 1/0.1`.

$Var(X) = \frac{1-p}{p^2} = \frac{0.9}{(0.1)^2} =$ `r 0.9/(0.1)^2`.

Standard deviation = $\sqrt{Var(X)} = \sqrt{90} \approx$ `r sqrt(90)`.

```{r part-a}
pdf <- pgeom(8, 0.1, lower.tail = F)
p <- 0.1
q <- 1 - p
ex <- p^-1
var <- q/p^2
sd <- sqrt(var)
cat(sprintf("\n %s = %f \n", 
            c("Probability", "Expected Value", "Variance", "Standard Deviation"),
            c(pdf, ex, var, sd))
    )
```

### b.
What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential. 

#### Solution
$P(X > 8) = e^{-\lambda i}$, with $\lambda = 0.1$.

$P(X > 8) = e^{-0.8} =$ `r exp(-0.8)`.

$E[X] = \frac{1}{\lambda} =$ `r 1/0.1`.

$Var(X) = \frac{1}{\lambda^2} = \frac{1}{0.1^2} =$ `r 1/(0.1)^2`.

Standard Deviation $= \sqrt{Var(X)} = \sqrt{100} = 10$.

```{r part-b}
pdf <- pexp(8, 0.1, lower.tail = F)
l <- 0.1
ex <- 1/l
var <- 1/l^2
sd <- sqrt(var)
cat(sprintf("\n %s = %f \n", 
            c("Probability", "Expected Value", "Variance", "Standard Deviation"),
            c(pdf, ex, var, sd))
    )
```

### c.
What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)

#### Solution
We're looking for 0 successes in 8 years. So, $P(0)$ with $n = 8$.

$P(0) = \binom{8}{0}(0.1)^0 \times 0.9{8-0} = (0.9)^8 =$ `r (0.9)^8`.

$E[X] = np = 8\cdot 0.1 = 0.8$.

$Var(X) = npq = 0.8\cdot 0.9 =$ `r 0.8*0.9`.

Standard Deviation $= \sqrt{Var(X)} = \sqrt{0.72} =$ `r sqrt(0.72)`.

```{r part-c}
pdf <- pbinom(0, 8, 0.1)
n <- 8
i <- 0
p <- 0.1
q <- 0.9
ex <- n*p
var <- n*p*q
sd <- sqrt(var)
cat(sprintf("\n %s = %f \n", 
            c("Probability", "Expected Value", "Variance", "Standard Deviation"),
            c(pdf, ex, var, sd))
    )
```

### d.
What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson. 

#### Solution
Since Poisson uses averages, and we expect one failure every 10 years, we can say the average yearly failure rate is 0.1.
We're looking for 0 failures in the first 8 years.

$P_{i}(t) = \frac{(\lambda t)^{i}e^{-\lambda t}}{i!}$

$P_{0}(8) = \frac{(0.1\cdot 8)^{0}e^{-0.1\cdot 8}}{0!} = e^{-0.8} =$ `r exp(-0.8)`.

$E[X] = Var(X) = \lambda\cdot t = 0.1 \cdot 8 = 0.8$.

Standard Deviation $= \sqrt{Var(X)} = \sqrt{0.8} =$ `r sqrt(0.8)`.

```{r}
lambda <- 0.1
t <- 8
i <- 0
ex <- lambda*t
var <- lambda*t
sd <- sqrt(var)
pdf <- ppois(i, t*lambda)
cat(sprintf("\n %s = %f \n", 
            c("Probability", "Expected Value", "Variance", "Standard Deviation"),
            c(pdf, ex, var, sd))
    )
```