---
title: "DATA 605 - Homework 8"
author: "Omer Ozeren"
output:
  word_document:
    toc: yes
    toc_depth: '3'
  html_document:
    highlight: tango
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 7.2 Exercise 11

A company buys 100 lightbulbs, each of which has an exponential lifetime of
1000 hours. What is the expected time for the first of these bulbs to burn
out?

### SOLUTION
ANS: let x1,x2...xn be independent exponential random variable.

$$ X _i = \mu. e^{-\mu x}  where x \geq 0 $$
$$P(X>x) = e^{-\mu x} $$
$$ so, E(X) = 1/\ { \sum \limits_{1}^{n} \mu i } $$

X = first of these bulbs to burn out

n=100

$$\mu i=1/1000$$

i =1,2,...100

So, E(X) = 10 HRS.

## Chapter 7.2 Exercise 14

Assume that $X_1$ and $X_2$ are independent random variables, each having an
exponential density with parameter $\lambda$. Show that $Z = X_1-X_2$ has density
$f_Z(z) = (1/2)e^{-\lambda|z|}$.

#### SOLUTION

$f_Z(z) = (1/2)e^{-\lambda|z|}$ can be re-written as $f_Z(z) = \begin{cases} (1/2)e^{-\lambda z}, & \mbox{if } z \ge 0, \\ (1/2)e^{\lambda z}, & \mbox{if }z <0. \end{cases}$

Since $X_1$ and $X_2$ have exponential density, their PDF is

<CENTER>
$f_{X_1}(x)=f_{X_2}(x)=\begin{cases} \lambda e^{-\lambda x}, & \mbox{if } x\ge 0, \\ 0, & \mbox{otherwise. }\end{cases}$
</CENTER>

$$
\begin{split}
  f_Z(z) &= f_{X_1+(-X_2)}(z) \\
    &= \int_{-\infty}^{\infty} f_{-X_2}(z-x_1) f_{X_1}(x_1) dx_1 \\
    &= \int_{-\infty}^{\infty} f_{X_2}(x_1-z) f_{X_1}(x_1) dx_1 \\
    &= \int_{-\infty}^{\infty} \lambda e^{-\lambda(x_1-z)} \lambda e^{-\lambda x_1} dx_1 \\
    &= \int_{-\infty}^{\infty} \lambda^2 e^{-\lambda x_1 + \lambda z} e^{-\lambda x_1} dx_1 \\
    &= \int_{-\infty}^{\infty} \lambda^2 e^{\lambda z - \lambda x_1 - \lambda x_1} dx_1 \\
    &= \int_{-\infty}^{\infty} \lambda^2 e^{\lambda(z-2x_1)} dx_1
\end{split}
$$

Consider $z=x_1-x_2$, then $x_2=x_1-z$.

If $z \ge 0$, then $x_2=(x_1-z) \ge 0$, and $x_1 \ge z$, and, using WolframAlpha, $f_Z(z) = \int_{z}^{\infty} \lambda^2 e^{\lambda(z-2x_1)} dx_1 = \frac{1}{2} \lambda e^{-\lambda z}$. 

If $z < 0$, then $x_2=(x_1-z) \ge 0$, and $x_1 \ge 0$, and $f_Z(z) = \int_{0}^{\infty} \lambda^2 e^{\lambda(z-2x_1)} dx_1 =\frac{1}{2} \lambda e^{\lambda z}$.

Combining two sides we get $f_Z(z) = \begin{cases} (1/2)e^{-\lambda z}, & \mbox{if } z \ge 0, \\ (1/2)e^{\lambda z}, & \mbox{if }z <0. \end{cases}$

## Chapter 8.2 Exercise 1

Let $X$ be a continuous random variable with mean $\mu = 10$ and variance
$\sigma^2 = 100/3$. Using Chebyshev's Inequality, find an upper bound for the
following probabilities.

a. $P(|X - 10| \ge 2)$
b. $P(|X - 10| \ge 5)$
c. $P(|X - 10| \ge 9)$
d. $P(|X - 10| \ge 20)$

#### SOLUTION

Chebyshev Inequality: $P(|X-\mu|\ge\epsilon) \le \frac{\sigma^2}{\epsilon^2}$ .

Per problem, $\mu=10$ and $\sigma = \sqrt{\frac{100}{3}} = \frac{10}{\sqrt{3}}$.

If $\epsilon = k\sigma$, then $k=\frac{\epsilon}{\sigma} = \frac{\epsilon\sqrt{3}}{10}$.


#### (a) P(|X - 10|>=2).

Let us find k first


```{r a, include=TRUE}
s = 10/sqrt(3)
#sigma
s
#value of k
k = 2/s
k
#value of k square -- upper bound
ksq = 1/(k*k)
ksq
```

S0, P(|X - 10|>=2) <= 8.33 which is not true.


#### (b) P(|X - 10| >= 5).

```{r b, include=TRUE}
s = 10/sqrt(3)
#sigma
s
#value of k
k = 5/s
k
#value of k square  -- upper bound
ksq = 1/(k*k)
ksq
```


So, P(|X - 10|>=5) <= 1.33 which is not true


#### (c) P(|X - 10| >= 9).

```{r c, include=TRUE}
s = 10/sqrt(3)
#sigma
s
#value of k
k = 9/s
k
#value of k square  -- upper bound
ksq = 1/(k*k)
ksq
```



So, P(|X - 10|>=9) <= 0.4115 which is not true


#### (d) P(|X - 10| >= 20).


```{r d, include=TRUE}
s = 10/sqrt(3)
#sigma
s
#value of k
k = 20/s
k
#value of k square  -- upper bound
ksq = 1/(k*k)
ksq
```


So, P(|X - 10|>=20) <= 0.083 which is not true.