---
title: "DATA 605 - Discussion 3"
author: "Omer Ozeren"
output:
  html_document:
    highlight: tango
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '6'
---


### Chapter E Section EE Exercise C24 -Page 388

Find the eigenvalues, eigenspaces, algebraic and geometric multiplicities for 

$$
A = 
\begin{bmatrix}
  1 & -1 & 1 \\
  -1 & 1 & -1 \\
  1 & -1 & 1
\end{bmatrix}
$$

### Solution

```{r}
A <- matrix(c(1,-1,1,-1,1,-1,1,-1,1), nrow=3, byrow=TRUE)
A
```

#### **Eigenvalues**

```{r}
# Calculate characteristic polynomial using pracma library
library(pracma)
charpoly(A)
```

$p_A(x) = x^3-3x^2+0x+0 = x^2(x-3)$

$p_A(\lambda) = 0$, so **eigenvalues** are $\lambda = 0$ and $\lambda = 3$.

#### **Algebraic multiplicities**

$\alpha_A(0) = 2$ and $\alpha_A(3) = 1$

#### **Eigenspace**

If $\lambda=0$, then $A - 0I_3 = A$ and it is row-reduced to 

```{r}
rref(A)
```

Then **eigenspace** is  
$$
 \Bigg\langle \Bigg\{
 \begin{bmatrix}
 1\\
 1\\
 0
 \end{bmatrix}
 ,
 \begin{bmatrix}
 -1\\
 0\\
 1
 \end{bmatrix}
 \Bigg\} \Bigg \rangle
$$

If $\lambda=3$, then $A - 3I_3$ is row-reduced to 

```{r}
rref(A - 3 * diag(3))
```

Then **eigenspace** is  
$$
 \Bigg\langle \Bigg\{
 \begin{bmatrix}
 1\\
 -1\\
 1
 \end{bmatrix}
 \Bigg\} \Bigg \rangle
$$

#### **Geometric multiplicities**

$\gamma_A(0) = 2$ and $\gamma_A(3) = 1$

### Notes

`eigen` function again returns a very small, near-zero value instead of $0$.

```{r}
eigen(A)$values
```

Additionally, per theorem DMFE since $\alpha_A(\lambda) = \gamma_A(\lambda)$ for all $\lambda$, then $A$ should be diagonalizable and eigenvalues should be on the diagonal. Consider $S$ consisting of eigenvectors.

```{r}
S <- matrix(c(1,-1,1,1,0,-1,0,1,1), nrow=3, byrow=TRUE)
S
```

Compute $S^{-1}AS$

```{r}
inv(S) %*% A %*% S
```

Not considering near-zero values, everything checks out.