---
title: "DATA 605 - Discussion 11"
author: "Omer Ozeren"
output:
  word_document:
    toc: yes
    toc_depth: '3'
  html_document:
    highlight: tango
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

    Using R, build a  regression model for data that interests you.  Conduct residual analysis.
    Was the linear model appropriate? Why or why not?

### For this discussion, I will look at Kaggle's Powerlifting Database dataset.

- It's a dataset containing competitor results in powerlifting from the OpenPowerlifting Database and do residual analysis.

  I will build a simple linear regression model of body weights vs best bench press for seniors to see if a linear relation exists between them.

- Dataset can be found here: https://www.kaggle.com/open-powerlifting/powerlifting-database

- Get the data and examine a preview

```{r}
library(data.table)
powerlift <- read.csv('C:\\Users\\OMERO\\Documents\\GitHub\\DATA605\\openpowerlifting.csv')
head(powerlift, n=5)
```

```{r, message=FALSE}
library(dplyr)
powerlift_senior <- powerlift %>% filter(Best3BenchKg > 0 & Division == 'Senior') %>% 
  select(BodyweightKg, Best3BenchKg)
```

### Visualize the data (EDA)

```{r, message=FALSE}
with(powerlift_senior, plot(BodyweightKg, Best3BenchKg, xlab = "Body Weight (Kg)",
                            ylab = "Best Bench Press (Kg)"))
```

### Residual Analysis

```{r}
lm_powerlift_senior <- lm(Best3BenchKg ~ BodyweightKg, data = powerlift_senior)
with(powerlift_senior, plot(BodyweightKg, Best3BenchKg,xlab = "Body Weight (Kg)",
                            ylab = "Best Bench Press (Kg)"))
abline(lm_powerlift_senior)
summary(lm_powerlift_senior)
```

- Equation of line is

$$
\begin{aligned}
\widehat{bestbenchpress} = 0.1376 + 1.2804*\widehat{bodyweight}
\end{aligned}
$$
#### Residual plots
```{r}
plot(fitted(lm_powerlift_senior), resid(lm_powerlift_senior),
     main = "Fitted vs residuals", xlab = "", ylab = "")
abline(h =0)
hist(resid(lm_powerlift_senior), xlab = "", main = "Histogram of Residuals")
qqnorm(resid(lm_powerlift_senior))
qqline(resid(lm_powerlift_senior))
```
### Summary

- We see that a linear model based on one explanatory variable doesn't explain the

data well. The $R^2$ value is quite low which shows that the fitted model doesn't

accuractely predict the values of Senior divisions competitiors bench press best based 

on their weight.