---
title: "DATA 605 - Discussion 5"
author: "Omer Ozeren"
output:
  html_document:
    highlight: tango
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '5'
---

## Chapter 1 Exercise 3 :

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```
**In the early 1600s, Galileo was asked to explain the fact that, although the**
**number of triples of integers from 1 to 6 with sum 9 is the same as the number**
**of such triples with sum 10, when three dice are rolled, a 9 seemed to come**
**up less often than a 10—supposedly in the experience of gamblers.**
 
 **(a) Write a program to simulate the roll of three dice a large number of**
**times and keep track of the proportion of times that the sum is 9 and**
**the proportion of times it is 10.**

**(b) Can you conclude from your simulations that the gamblers were correct?**


## Solutions :

### (a)

```{r}
# number of times to run this simulation of 10000 trials/runs
n_sim <- 100 # 100 times simulate 10000 trials
total_times <- rep(FALSE, n_sim) # vector to store all simulation results
for (i in 1:n_sim){
  
  sum_9 <- 0 # counter for # of times sum == 9
  sum_10 <- 0 # counter for # of times sum == 10
  n <- 10000 # number of times to roll 3 dice
  # for loop to simulate 
  for (j in 1:n){
    dice <- sample(1:6, 3, replace = TRUE) # pick a number from 1 to 6 3 times with equal likelihood
    if (sum(dice) == 9){ # if the sum of the rolled 3 dice is 9
      sum_9 <- sum_9 + 1 # increment the counter
    }
    if (sum(dice) == 10){ # if the sum of the rolled 3 dice is 10
      sum_10 <- sum_10 + 1
    }
  }
  
  # check if the # of sum == 9 is less than # of sum == 10
  if (sum_9 < sum_10){
    total_times[i] <- TRUE # mark as true
  }
}
```

**Running this simulation 100 times of rolling 3 dice 10000 times:**

```{r}
print(sum(total_times == TRUE))
```

### (b)

**The gamblers were right. The sum of 9 comes up less frequently than sum of 10 for rolling 3 dice usually around 90% or more of the time.**