---
title: "DATA 605 - Final Exam"
author: "Omer Ozeren"
output:
  html_document:
    highlight: tango
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T)
```
## Problem 1 :

Using R, generate a random variable X that has 10,000 random uniform numbers from 1 to N, where N can be any number of your choosing greater than or equal to 6.  Then generate a random variable Y that has 10,000 random normal numbers with a mean of mean=std=(N+1)/2

```{r data}
#10,000 random uniform numbers from 1 to 7
N=20
# 10,000 random uniform numbers from 1 to N
X = runif(1000, 1,N)
# 10,000 random normal numbers with a mean of mean=std=(N+1)/2
Y = rnorm(1000, (N+1)/2, (N+1)/2)
```
##Probability

Calculate as a minimum the below probabilities a through c.  Assume the small letter "x" is estimated as the median of the X variable, and the small letter "y" is estimated as the 1st quartile of the Y variable.  Interpret the meaning of all probabilities

```{r}
XY<- cbind(X,Y)
var <- nrow(XY)
x <- median(X)
y <- quantile(Y, 0.25,names=FALSE)
```

### A: $$P(X>x|X>y)$$
\[
P(X > x | X > y) = \frac{P(X > x \ and \ X > y)}{P(X > y)}
\]
```{r}
XGy <- length(which(X>y))
XGy_XGx <- length(which(X>y & X>x))
XGy_XGx/XGy
```


### B: $$P(X>x , Y>y)$$

We know the statistics of half of the values in X are above the median, and 75% of the values in Y are above the first quartile
\[
P(X > x, Y > Y) = P(X > x \ and \ Y > y)
\]

\[
P(X > x) = 0.5
\]

\[
P(Y > y) = 0.75
\]

\[
P(X > x \ and \ Y > y) = (0.5)(0.75) =0.375
\]

### C: $$P(X<x|Y>y)$$

```{r}
XGy <- length(which(X>y))
XGy_xGX <- length(which(X>y & X<x))

XGy_xGX/XGy
```
