---
title: "DATA 605 - Homework 14"
author: "Omer Ozeren"
output:
  word_document:
    toc: yes
    toc_depth: '3'
  html_document:
    highlight: tango
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T)
```

   This week, we'll work out some Taylor Series expansions of popular functions.
    
- $f(x) = 1/(1-x)$
- $f(x) = e^x$
- $f(x) = \ln(1+x)$

    For each function, only consider its valid ranges as indicated in the notes when
    you are computing the Taylor Series expansion. Please submit your assignment as a
    R-Markdown document.
    
## Function 1: $f(x) = 1/(1-x)$

$$f(x) =\frac {1} {(1-x)} $$
$$f'(x) = \frac {1}{(1-x)^2} $$

$$f''(x) = \frac {2}{(1-x)^3} $$

$$f'''(x) = \frac {6}{(1-x)^4} $$

$$ = f(x) + \frac {f'(x)} {1!} (x -c)^1 + \frac {f''(x)} {2!} (x-c)^2 + \frac {f'''(x)} {3!} (x -c)^3$$ 

$$ = 1 + x + x^2 + x^3 + x^4 + ...$$


$$  \sum _{n=0}^{\infty} x^n $$
```{r c1}
library(pracma)
c = -5
f <- function(x) {1/(1-x)} 
A1 <- taylor(f, x0=c, 5)
A1
```    


## Function 2: $f(x) = e^x$

Find first several derivatives.

$f^0(c) = e^c$

$f'(c) = e^c$

$f''(c) = e^c$

$f'''(c) = e^c$

$f''''(c) = e^c$

Per definition, 

$$
\begin{split}
f(x) &= \frac{e^c}{0!}(x-c)^0 + \frac{e^c}{1!}(x-c)^1 + \frac{e^c}{2!}(x-c)^2 + \frac{e^c}{3!}(x-c)^3 + ...\\
&= e^c + e^c(x-c) + e^c\frac{(x-c)^2}{2!} + e^c\frac{(x-c)^3}{3!} + ...\\
&= e^c \sum\limits_{n=0}^{\infty} \frac{(x-c)^n}{n!}
\end{split}
$$

The Maclaurin Series of $f(x)$, $c=0$, $f(x) = \sum\limits_{n=0}^{\infty} \frac{x^n}{n!} = 1 + x + \frac{x^2}{2} + \frac{x^3}{6} + \frac{x^4}{24} + ...$.

```{r c2}
c = -5
f <- function(x) {exp(x)} 
A2 <- taylor(f, x0=c, 5)
A2

```
## Function 3: $f(x) = ln(1+x)$ 

$$f(x) = ln(1 + x)$$ 
$$f'(x) = \frac {1}{1+x} $$
$$f''(x) = \frac {-1}{(1+x)^2} $$
$$f'''(x) = \frac {2}{(1+x)^3} $$
$$f(x) + \frac {f'(x)} {1!} x^1 + \frac {f''(x)} {2!} x^2 $$ 
$$ f(x) = f(0) +f'(0)(x - c) + \frac {f''(c)}{2!} (x-c)^2 + \frac {f'''(c)}{3!} (x- c)^3 + $$
$So the Taylor series for f(x)=ln(1+x) is:$
$$ f(x) = x - \frac {1} {2} x^2 + \frac {1}{3} x^3 - \frac {1}{4} x^4 + ... $$
$$  \sum _{n=1}^{\infty} (-1)^{n+1} * \frac{x^n}{n} $$
```{r c3}
c = 0
f <- function(x) {log(1+x)} 
A3 <- taylor(f, x0=c, 5)
A3
```